---
title: "Male Neuroplasticity Study: Brain Structure Volume Analysis for Male Bombus Brains"
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE)
```

***
                  
                           
```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(car)
library(stats)
library(nortest)
library(MASS)

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
\newpage


```{r loaddata, include=FALSE}
#read in data
KE_maleNP <- read.csv2("Data/Bimpatiens_MNP_StructureVolumes.csv", sep=",", stringsAsFactors = FALSE)



#remove extra rows
KE_maleNP <- KE_maleNP[1:18,] #make sure no extra rows
print(str(KE_maleNP))

KE_maleNP <- KE_maleNP[,-c(9:10,17:24)] #remove total columns KE had; remove proportions already calculated by KE
#Calculations will be done below to ensure repeatability and to remain consistent across data sets
```

```{r DATASTUFF, echo=FALSE}
#Make TX a factor
KE_maleNP$treatment <- factor(KE_maleNP$treatment, levels = c("NE", "AGED"))

#make proportions type numeric
KE_maleNP$mb_lobe<- as.numeric(KE_maleNP$mb_lobe) #MBLobe
KE_maleNP$antennal_lobe <- as.numeric(KE_maleNP$antennal_lobe) #AL
KE_maleNP$lateral_kenyoncells <- as.numeric(KE_maleNP$lateral_kenyoncells) #Lateral KC
KE_maleNP$medial_kenyoncells <- as.numeric(KE_maleNP$medial_kenyoncells) #medial KC
KE_maleNP$lateral_calyx <- as.numeric(KE_maleNP$lateral_calyx) #lateralcalyx
KE_maleNP$medial_calyx <- as.numeric(KE_maleNP$medial_calyx) #medialcalyx

#wholebrain
KE_maleNP$whole_brain <- as.numeric(KE_maleNP$whole_brain) #whole brain

print(str(KE_maleNP))

#Reordering formatting
NE <- subset(KE_maleNP, treatment=='NE')
AGED <- subset(KE_maleNP, treatment=='AGED')

KE_NPdata <- rbind(NE,AGED)
```


\section{Check for duplicate values}
```{r duplicatevalues}
anyDuplicated(KE_NPdata[,8:14])
#no duplicate values
```




\section{Combine Lateral and Medial Measurements for Select Structures}
\subsection{Summing Measurements}
Before analyzing the data we need to sum medial and lateral measurements to get one value for each structure (i.e. ```lateral_calyx + medial_calyx``` to get one value for the calyces.)

The table below summarizes how the values were summed and their new naming code.
\subsection{Record of what data are being summed}
```{r collapsedtable, echo=FALSE}
#library(knitr)
sumstructures <- c("calyces","mushroom body lobe", "kenyon cells", "neuropil")

sumstructcode <- c("calyces", "mblobe", "kcs", "neuropil")

collapsed_structures <- c( "lateral_calyx  +  medial_calyx", "mblobe", "lateral_kc  +  medial_kc", "sum all neuropil for N:K")
sum_data_table <- cbind(sumstructures, sumstructcode, collapsed_structures)
colnms <- c("Summed Structures", "Name", "Individual Structures Collapsed")
kable(sum_data_table, col.names = colnms, align=c('l', 'l', 'l'), caption = "Information for specific structures that were summed to make 'combined' structures")
```


\subsection{Summing the data}
Below is an example of how data are summed across lateral and medial structures.
```{r example, eval=FALSE, echo=TRUE}
calyx <- apply(data[,c("lateral_calyx", "medial_calyx")], 1, sum)
```


```{r combinedata, echo=FALSE}
#to sum the calyces 
calyces <- apply(KE_NPdata[,c("lateral_calyx","medial_calyx")], 1, sum)

#to sum the kenyon cells 
kcs <- apply(KE_NPdata[,c("lateral_kenyoncells", "medial_kenyoncells")], 1, sum)

#to sum the neuropil 
neuropil <- apply(KE_NPdata[,c("lateral_calyx", "medial_calyx", "mb_lobe")], 1, sum)

#bind to create new dataframe containing these columns
Bimp_maleNP_plussumdata <- cbind(KE_NPdata, calyces, kcs, neuropil)
```


```{r sumstructure, echo=FALSE}
#Save summed data
write.csv(Bimp_maleNP_plussumdata, "Data/Bimp_MNP_plussummeddata.csv")
```


\section{Summed Structures Scaled Relative to Whole Brain}
Structures that were summed in the prior step are now made ratios of whole brain measurements.

\subsection{Make New Table: Containing Summed Structures}
```{r mk_tableonlysummedstructures}
#make a new table with only the factors and the data we want to take whole brain proportions of...
Bimp_maleNP_sumdata <- Bimp_maleNP_plussumdata[,-c(9:12)]
```

\subsection{Scale to whole brain}
```{r relativetoWB}
#calculate relative to wholebrain
#using mutate we can add the calculated column to our dataframe

#library(dplyr)
#relative to Whole brain
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, calyces_relWB = calyces/whole_brain) # calyces
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, mblobe_relWB = mb_lobe/whole_brain) # mblobes
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, kcs_relWB = kcs/whole_brain) # kcs
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, neuropil_relWB = neuropil/whole_brain) # neuropil

#N:K
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, nk = neuropil/kcs) # N:K

#relative to KCS
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, calyces_relKC = calyces/kcs) # calyces
Bimp_maleNP_sumdata <- mutate(Bimp_maleNP_sumdata, mblobe_relKC = mb_lobe/kcs) # mblobes
```



```{r saveprop, echo=FALSE}
#Save Prop Data
write.csv(Bimp_maleNP_sumdata, "Data/Bimp_MNP_proportiondata.csv")

```


\subsection{Make New Table: just the proportion data}
```{r mk_propdata}
#make a new table with only the factors whole brain proportion data
KE_NPdata <- Bimp_maleNP_sumdata[,-c(8:9,11,13)] #KEEP raw Wholebrain and KCS for t-test
```




\section{Whole Brain t-Test}
\subsection{Summary statistics}
```{r ss_wb, echo=FALSE}
#library(Rmisc)
summarystats_wb <- summarySE(KE_NPdata,
                                 measurevar = "whole_brain",
                                 groupvars = c("treatment"))


kable(summarystats_wb)
```


```{r bp_wb, echo=FALSE}
#outlier function
is_outlier <- function(x){
  return(x < quantile(x, 0.25)-1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

samples_KE_NPdata<- KE_NPdata[,-1]
rownames(samples_KE_NPdata) <- KE_NPdata[,1]

outlier_test_whole_brain <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(whole_brain), whole_brain, as.numeric(NA)))

outlier_test_whole_brain$outlier[which(is.na(outlier_test_whole_brain$is_outlier))] <- as.numeric(NA)




#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=whole_brain, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Whole Brain") +
  annotate("text", x=2, y=175000000, label="K18R.m017", size=4, col="red") +
  scale_y_continuous(name = "Volume", limits = c(90000000,190000000)) +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```


\subsection{T-test}
```{r t-test_wb, include=TRUE}
tt_wb <- t.test(whole_brain ~ treatment, data = KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_wb)

tt_wb$statistic
sprintf("%.10f", tt_wb$p.value)


#Calculate percent increase between means
p.inc <- function(meanAged,meanNE){
 return((meanAged-meanNE)/(meanNE) * 100)
}

#calculate % increase
wb_PI <- (p.inc(tt_wb$estimate[[2]], tt_wb$estimate[[1]]))
wb_PI

```


Whole brain mean volume of AGED males (`r tt_wb$estimate[[2]]`) is `r round(wb_PI,2)`% greater than NE males (`r tt_wb$estimate[[1]]`).



\subsection{Checking for Normality and Homogeneity of Variance}
```{r testassumption_wb}
qqPlot(KE_NPdata$whole_brain, "norm")
print(ad.test(KE_NPdata$whole_brain))
print(leveneTest(whole_brain~treatment, data = KE_NPdata))
```





\section{Kenyon cells t-Test}
\subsection{Summary statistics}
```{r ss_kcsraw, echo=FALSE}
#library(Rmisc)
summarystats_kcsraw <- summarySE(KE_NPdata,
                                 measurevar = "kcs",
                                 groupvars = c("treatment"))


kable(summarystats_kcsraw)
```


```{r bp_kcsraw, echo=FALSE}
outlier_test_kcsraw <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(kcs), kcs, as.numeric(NA)))

outlier_test_kcsraw$outlier[which(is.na(outlier_test_kcsraw$is_outlier))] <- as.numeric(NA)




#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=kcs, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Kenyon Cells: Not standardized") +
  scale_y_continuous(name = "Volume", limits = c(3000000,12000000)) +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```


\subsection{T-test}
```{r t-test_kcsraw, include=TRUE}
tt_kcsraw <- t.test(kcs ~ treatment, data = KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_kcsraw)

tt_kcsraw$statistic
sprintf("%.10f", tt_kcsraw$p.value)
```


\subsection{Checking for Normality and Homogeneity of Variance}
```{r testassumption_kcsraw}
qqPlot(KE_NPdata$kcs, "norm")
print(ad.test(KE_NPdata$kcs))
print(leveneTest(kcs~treatment, data = KE_NPdata))
```









\section{Data Anaylsis: T-tests comparing relative volumes (whole brain) of brain structures between Newly-Emerged and Aged bees}

\subsection{Calyces: lip, collar, basal ring}
\subsubsection{Summary statistics}
```{r ss_calyce, echo=FALSE}
#library(Rmisc)
summarystats_calyces <- summarySE(KE_NPdata,
                                 measurevar = "calyces_relWB",
                                 groupvars = c("treatment"))


kable(summarystats_calyces)
```



```{r bp_cal, echo=FALSE}
outlier_test_calyces <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(calyces_relWB), calyces_relWB, as.numeric(NA)))

outlier_test_calyces$outlier[which(is.na(outlier_test_calyces$is_outlier))] <- as.numeric(NA)




#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=calyces_relWB, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Calyces") +
  annotate("text", x=1, y=0.085, label="K18R.m010", size=4, col="red") +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```


\subsubsection{T-test}
```{r t-test_cal, include=TRUE}
tt_calyces <- t.test(calyces_relWB~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_calyces)

tt_calyces$statistic
sprintf("%.10f", tt_calyces$p.value)


#calculate % increase
cal_PI <- (p.inc(tt_calyces$estimate[[2]], tt_calyces$estimate[[1]]))
cal_PI
```

Calyx mean volume of AGED males (`r tt_calyces$estimate[[2]]`) is `r round(cal_PI,2)`% greater than NE males (`r tt_calyces$estimate[[1]]`).


\subsubsection{Checking for Normality and Homogeneity of Variance}
```{r t-test_assump}
qqPlot(KE_NPdata$calyces_relWB, "norm")
print(ad.test(KE_NPdata$calyces_relWB))
print(leveneTest(calyces_relWB~treatment, data = KE_NPdata))
```

Normal based on Q-Q plot but not the other tests. So, going to box cox transform.


\subsubsection{Box-Cox Transformation}
```{r BoxCox_calyces, echo=TRUE, eval=TRUE, fig.height=3.5}
##Box-Cox Transformation
bc_calyces <- boxcox(calyces_relWB ~ treatment, 
                       data = KE_NPdata, lambda= seq(-20, 14, length=100), plotit = TRUE)

#get lambda
lambda_calyces <- bc_calyces$x
#get log-likelyhood
likelyhood_calyces <- bc_calyces$y

bcresults_calyces <- cbind(lambda_calyces, likelyhood_calyces)

#sort these in decending order
ord_bc_calyces <-bcresults_calyces[order(-likelyhood_calyces),]

#pull out maximum likelyhood and corresponding lambda
max_lik_calyces<- ord_bc_calyces[1,]
#lambda_calyces likelyhood_calyces

new.lambda_calyces <- ord_bc_calyces[1,1]
```

Using the ```boxcox()``` function from the package ```MASS```, we determined that $\lambda$ =  `r new.lambda_calyces`.



\subsubsection{Box-cox transformation model}
```{r transformed_calycesboxcoxtt}
box_ttest_calyces <- t.test((((KE_NPdata$calyces_relWB ^ 
                new.lambda_calyces)-1)/new.lambda_calyces)~treatment, 
                data=KE_NPdata, alternative=c("two.sided"), paired=FALSE, 
                var.equal=TRUE)
  
print(box_ttest_calyces)

#get full for MS
box_ttest_calyces$statistic
sprintf("%.10f", box_ttest_calyces$p.value)



cal_bc_PI <- (p.inc(box_ttest_calyces$estimate[[2]], box_ttest_calyces$estimate[[1]]))
cal_bc_PI
```

Calyx mean volume of AGED males (`r box_ttest_calyces$estimate[[2]]`) is `r round(cal_bc_PI,2)`% greater than NE males (`r box_ttest_calyces$estimate[[1]]`).


\subsubsection{Retest Assumptions}
```{r transformed_assumptions, echo=FALSE}
qqPlot((((KE_NPdata$calyces_relWB ^ 
                new.lambda_calyces)-1)/new.lambda_calyces), "norm")


print(ad.test((((KE_NPdata$calyces_relWB ^ 
                new.lambda_calyces)-1)/new.lambda_calyces)))

print(leveneTest((((KE_NPdata$calyces_relWB ^ 
                new.lambda_calyces)-1)/new.lambda_calyces) ~ treatment, 
                          data = KE_NPdata))
```





\subsection{Kenyon Cells: Total Kenyon Cell}
\subsubsection{summary statistics}
```{r ss_kc, echo=FALSE}
#library(Rmisc)
summarystats_kc <- summarySE(KE_NPdata,
                             measurevar = "kcs_relWB",
                             groupvars = c("treatment"))


kable(summarystats_kc)

```

```{r bp_kc, echo=FALSE}
#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=kcs_relWB, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Kenyon cells") +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```


\subsubsection{T-test}
```{r t-test_kc}
tt_kcs <- t.test(kcs_relWB~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_kcs)

tt_kcs$statistic
sprintf("%.10f", tt_kcs$p.value)

```




\subsubsection{Checking for Normality and Homogeneity of Variance}
```{r t-test_assump_kc}
qqPlot(KE_NPdata$kcs_relWB, "norm")
print(ad.test(KE_NPdata$kcs_relWB))
print(leveneTest(kcs_relWB~treatment, data = KE_NPdata))
```



\subsubsection{Box-Cox Transformation}
```{r BoxCox_kcs, echo=TRUE, eval=TRUE, fig.height=3.5}
##Box-Cox Transformation
bc_kcs <- boxcox(kcs_relWB ~ treatment, 
                       data = KE_NPdata, lambda= seq(-25, 14, length=100), plotit = TRUE)

#get lambda
lambda_kcs <- bc_kcs$x
#get log-likelyhood
likelyhood_kcs <- bc_kcs$y

bcresults_kcs <- cbind(lambda_kcs, likelyhood_kcs)

#sort these in decending order
ord_bc_kcs <-bcresults_kcs[order(-likelyhood_kcs),]

#pull out maximum likelyhood and corresponding lambda
max_lik_kcs <- ord_bc_kcs[1,]
#lambda_kcs likelyhood_kcs 

new.lambda_kcs <- ord_bc_kcs[1,1]
```

Using the ```boxcox()``` function from the package ```MASS```, we determined that $\lambda$ =  `r new.lambda_kcs`.



\subsubsection{Box-cox transformation model}
```{r transformed_kcsboxcoxtt}
box_ttest_kcs <- t.test((((KE_NPdata$kcs_relWB ^ 
                new.lambda_kcs)-1)/new.lambda_kcs) ~ treatment, 
                data=KE_NPdata, alternative=c("two.sided"), paired=FALSE, 
                var.equal=TRUE)
  
print(box_ttest_kcs)

#get full for MS
box_ttest_kcs$statistic
sprintf("%.10f", box_ttest_kcs$p.value)

```



\subsubsection{Retest Assumptions}
```{r transformed_assumptions_KC, echo=FALSE}
qqPlot((((KE_NPdata$kcs_relWB ^ 
                new.lambda_kcs)-1)/new.lambda_kcs), "norm")


print(ad.test((((KE_NPdata$kcs_relWB ^ 
                new.lambda_kcs)-1)/new.lambda_kcs)))

print(leveneTest((((KE_NPdata$kcs_relWB ^ 
                new.lambda_kcs)-1)/new.lambda_kcs) ~ treatment, 
                          data = KE_NPdata))
```










\subsection{Mushroom Body Lobes: Including Peduncle and Sensory Lobes}
\subsubsection{summary statistics}
```{r ss_mblobe, echo=FALSE}
#library(Rmisc)
summarystats_mblobe <- summarySE(KE_NPdata,
                                 measurevar = "mblobe_relWB",
                                 groupvars = c("treatment"))

kable(summarystats_mblobe)

```


```{r bp_mblobe, echo=FALSE}
#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=mblobe_relWB, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Mushroom Body Lobes") +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```


\subsubsection{T-test}
```{r t-test_mblobe, include=TRUE}
tt_mblobe <- t.test(mblobe_relWB~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_mblobe)

tt_mblobe$statistic
sprintf("%.10f", tt_mblobe$p.value)
```



\subsubsection{Checking for Normality and Homogeneity of Variance}
```{r t-test_assump_mblobe}
qqPlot(KE_NPdata$mblobe_relWB, "norm")
print(ad.test(KE_NPdata$mblobe_relWB))
print(leveneTest(mblobe_relWB~treatment, data = KE_NPdata))
```






\subsection{Neuropil}
\subsubsection{summary statistics}
```{r ss_N, echo=FALSE}
#library(Rmisc)
summarystats_N <- summarySE(KE_NPdata,
                                 measurevar = "neuropil_relWB",
                                 groupvars = c("treatment"))

kable(summarystats_N)
```

```{r bp_N, echo=FALSE}

outlier_test_neuropil <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(neuropil_relWB), neuropil_relWB, as.numeric(NA)))

outlier_test_neuropil$outlier[which(is.na(outlier_test_neuropil$is_outlier))] <- as.numeric(NA)


#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=neuropil_relWB, fill=treatment)) +
  geom_boxplot() +
  annotate("text", x=1, y=0.1335, label="K18R.m010", size=4, col="red") +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("MB Neuropil") +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))

```

\subsubsection{T-test}
```{r t-test_N, include=TRUE}

tt_neuropil <- t.test(neuropil_relWB~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_neuropil)

tt_neuropil$statistic
sprintf("%.10f", tt_neuropil$p.value)


#% increase
neuropil_PI <- (p.inc(tt_neuropil$estimate[[2]], tt_neuropil$estimate[[1]]))
neuropil_PI
```

Neuropil mean volume of AGED males (`r tt_neuropil$estimate[[2]]`) is `r round(neuropil_PI,2)`% greater than NE males (`r tt_neuropil$estimate[[1]]`).




\subsubsection{Checking for Normality and Homogeneity of Variance}
```{r t-test_assump_N}
qqPlot(KE_NPdata$neuropil_relWB, "norm")
print(ad.test(KE_NPdata$neuropil_relWB))
print(leveneTest(neuropil_relWB~treatment, data = KE_NPdata))
```







\subsection{N:K}
\subsubsection{summary statistics}
```{r ss_nk, echo=FALSE}
#library(Rmisc)
summarystats_nk <- summarySE(KE_NPdata,
                                 measurevar = "nk",
                                 groupvars = c("treatment"))

kable(summarystats_nk)
```

```{r bp_nk, echo=FALSE}

outlier_test_nk <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(nk), nk, as.numeric(NA)))

outlier_test_nk$outlier[which(is.na(outlier_test_nk$is_outlier))] <- as.numeric(NA)


#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=nk, fill=treatment)) +
  geom_boxplot() +
  annotate("text", x=2, y=2.9, label="K18P.m003", size=4, col="red") +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("N:K") +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))

```

\subsubsection{T-test}
```{r t-test_nk, include=TRUE}
tt_nk <- t.test(nk~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_nk)

tt_nk$statistic
sprintf("%.10f", tt_nk$p.value)

#% increase
nk_PI <- (p.inc(tt_nk$estimate[[2]], tt_nk$estimate[[1]]))
nk_PI
```

NK mean of AGED males (`r tt_nk$estimate[[2]]`) is `r round(nk_PI,2)`% greater than NE males (`r tt_nk$estimate[[1]]`).



\subsubsection{Checking for Normality and Homogeneity of Variance}
```{r t-test_assump_nk}
qqPlot(KE_NPdata$nk, "norm")
print(ad.test(KE_NPdata$nk))
print(leveneTest(nk~treatment, data = KE_NPdata))
```











\section{Data Anaylsis: Relative Volumes (Kenyon cells) Between Newly-Emerged and Aged bees}
\subsection{Calyces: lip, collar, basal ring}
\subsubsection{Summary Statistics}
```{r meansum_calyx_byKC, echo=FALSE}
#library(Rmisc)
kable(summarySE(KE_NPdata,
                          measurevar = "calyces_relKC",
                          groupvars = c("treatment")))
```


```{r outlier_calyx_byKC, echo=FALSE}
outlier_test_calyces_relKC <- samples_KE_NPdata %>% tibble::rownames_to_column(var="outlier") %>% group_by(treatment) %>% mutate(is_outlier=ifelse(is_outlier(calyces_relKC), calyces_relKC, as.numeric(NA)))

outlier_test_calyces_relKC$outlier[which(is.na(outlier_test_calyces_relKC$is_outlier))] <- as.numeric(NA)


#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=calyces_relKC, fill=treatment)) +
  geom_boxplot() +
  annotate("text", x=2, y=1.8, label="K18P.m003", size=4, col="red") +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Calyces") +
  scale_y_continuous(name = "Volume Relative to Kenyon Cell Bodies", limits = c(1,2)) +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```



\subsubsection{T-test}
```{r tt_calyx_byKC}
tt_calyces_byKC <- t.test(calyces_relKC~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_calyces_byKC)

tt_calyces_byKC$statistic
sprintf("%.10f", tt_calyces_byKC$p.value)

#% increase
calyces_byKC_PI <- (p.inc(tt_calyces_byKC$estimate[[2]], tt_calyces_byKC$estimate[[1]]))
calyces_byKC_PI
```

Calyces:KC mean volume of AGED males (`r tt_calyces_byKC$estimate[[2]]`) is `r round(calyces_byKC_PI,2)`% greater than NE males (`r tt_calyces_byKC$estimate[[1]]`).


\subsubsection{Verifying Assumptions}
Checking for:
<br />  
1) The data are normally distributed
<br />  
2) Homogeneity of Variance

```{r assumptions_calyx_byKC, echo=FALSE}
qqPlot(KE_NPdata$calyces_relKC, "norm")
print(ad.test(KE_NPdata$calyces_relKC))
print(leveneTest(calyces_relKC~treatment, data=KE_NPdata))
```

Normality: Assumption met.
<br />  
Variance: Assumption met.




\subsection{Mushroom Body Lobes}
\subsubsection{Summary Statistics}
```{r meansum_mblobe_byKC, echo=FALSE}
#library(Rmisc)
kable(summarySE(KE_NPdata,
                          measurevar = "mblobe_relKC",
                          groupvars = c("treatment")))
```


```{r bp_mblobe_byKC, echo=FALSE}
#Boxplot
ggplot(KE_NPdata, aes(x=treatment, y=mblobe_relKC, fill=treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("darkgoldenrod1", "cyan3")) +
  ggtitle("Mushroom Body Lobes") +
  scale_y_continuous(name = "Volume Relative to Kenyon Cells", limits = c(0.3,1)) +
  scale_x_discrete(name = "Treatment Group", labels = c("NE", "AGED")) +
  theme(text = element_text(color="black",size = 15),
        axis.title = element_text(color="black"), 
        axis.text.x = element_text(color="black", size = 15, margin = margin(t = 5, r =0, b = 20, l = 0)), 
        axis.text.y = element_text(color="black", size = 15, margin = margin(t = 0, r =5, b = 0, l = 20)),
        axis.ticks.length = unit (.5,"cm"),
        legend.position = 'none', 
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(size = 1, linetype = "solid",colour = "black"),
        plot.title = element_text(hjust=0.5))
```



\subsubsection{T-test}
```{r tt_mblobe_byKC}
tt_mblobe_byKC <- t.test(mblobe_relKC~treatment, data=KE_NPdata, 
                     alternative=c("two.sided"), paired=FALSE, var.equal=TRUE)

print(tt_mblobe_byKC)

tt_mblobe_byKC$statistic
sprintf("%.10f", tt_mblobe_byKC$p.value)



#% increase
mblobe_byKC_PI <- (p.inc(tt_mblobe_byKC$estimate[[2]], tt_mblobe_byKC$estimate[[1]]))
mblobe_byKC_PI
```

MB Lobe:KC mean volume of AGED males (`r tt_mblobe_byKC$estimate[[2]]`) is `r round(mblobe_byKC_PI,2)`% greater than NE males (`r tt_mblobe_byKC$estimate[[1]]`).

\subsubsection{Verifying Assumptions}
Checking for:
<br />  
1) The data are normally distributed
<br />  
2) Homogeneity of Variance

```{r assumptions_mblobe_byKC, echo=FALSE}
qqPlot(KE_NPdata$mblobe_relKC, "norm")

#normality
print(ad.test(KE_NPdata$mblobe_relKC))

#variance
print(leveneTest(mblobe_relKC~treatment, data=KE_NPdata))
```

Normality: Assumption met.
<br />  
Variance: Assumption met.









\newpage
\section{Summary Table of the Results}
```{r summarizeallresults, echo=FALSE}
structure <- c("Calyces", "Mushroom Body Lobe", "Kenyon Cells", "Neuropil", "NK")

ne_mean <- c(tt_calyces$estimate[1], tt_mblobe$estimate[1], tt_kcs$estimate[1], 
             tt_neuropil$estimate[1], tt_nk$estimate[1])

aged_mean <- c(tt_calyces$estimate[2], tt_mblobe$estimate[2], tt_kcs$estimate[2], 
               tt_neuropil$estimate[2], tt_nk$estimate[2])
  
CI_lower <- c(tt_calyces$conf.int[1], tt_mblobe$conf.int[1], tt_kcs$conf.int[1],
              tt_neuropil$estimate[1],tt_nk$conf.int[1])
  
CI_upper <- c(tt_calyces$conf.int[2], tt_mblobe$conf.int[2], tt_kcs$conf.int[2],
              tt_neuropil$estimate[2],tt_nk$conf.int[2])

p_value <-c(tt_calyces$p.value, tt_mblobe$p.value, tt_kcs$p.value, 
            tt_neuropil$p.value, tt_nk$p.value)



summary <-cbind(structure, signif(ne_mean, 3), signif(aged_mean, 3), signif(CI_lower, 3), signif(CI_upper, 3), signif(p_value,4))
  

kable(summary, row.names = FALSE, col.names = c("Structure", "NE means", "Aged means", "Lower 95% CI", "Upper 95% CI", "p-Value"), caption = "Results when scaled by whole brain volumes.")
```

\section{Summary Table of the Results}
```{r summarizeallresults_KCrelative, echo=FALSE}
structure_byKC <- c("Calyces", "Mushroom Body Lobe")

ne_mean_byKC <- c(tt_calyces_byKC$estimate[1], tt_mblobe_byKC$estimate[1])

aged_mean_byKC <- c(tt_calyces_byKC$estimate[2], tt_mblobe_byKC$estimate[2])
  
CI_lower_byKC <- c(tt_calyces_byKC$conf.int[1], tt_mblobe_byKC$conf.int[1])
  
CI_upper_byKC <- c(tt_calyces_byKC$conf.int[2], tt_mblobe_byKC$conf.int[2])

p_value_byKC <-c(tt_calyces_byKC$p.value, tt_mblobe_byKC$p.value)



summary <-cbind(structure_byKC, signif(ne_mean_byKC, 3), signif(aged_mean_byKC, 3), signif(CI_lower_byKC, 3), signif(CI_upper_byKC, 3), signif(p_value_byKC,4))
  

kable(summary, row.names = FALSE, col.names = c("Structure", "NE means", "Aged means", "Lower 95% CI", "Upper 95% CI", "p-Value"), caption = "Results for MB neuropil when scaled by Kenyon cell volumes.")
```


